\documentclass[11pt]{article}

\usepackage{preamble}

\title{Cohomology in AC/DC RLC Circuits and Measurement}
\author{Colin Roberts}
\date{}

\begin{document}

\maketitle

\begin{abstract}
We present a cohomological viewpoint on lumped AC/DC RLC circuits, organized around a two-layer separation: topology determines which state spaces and pairings exist, while geometry and constitutive data (via a Hodge/constitutive operator and energy functionals) determine numerical parameters such as $R$, $L$, and $C$.
At the lumped level, Kirchhoff constraints are expressed using chains/cochains on a graph, with the first cohomology measuring obstructions to globally defined potentials.
At the field-domain level, Alexander duality and Poincar\'e--Lefschetz duality identify capacitive and inductive degrees of freedom and interpret idealized measurement as evaluation pairings.
At the compositional level, excision and Mayer--Vietoris control how these state spaces behave under interconnection before any constitutive law is chosen, while Dirichlet-to-Neumann maps and Schur complements recover the familiar effective parameters after constitutive data is added.
Periodic time ($S^1$) organizes AC steady states via K\"unneth-type decompositions and universal $i\omega$ phase factors.
\end{abstract}

\tableofcontents

\section{Introduction}
\label{sec:intro}
Electrical circuits are often taught as an exercise in solving linear equations, but there is a persistent geometric theme: conservation laws look like exactness, potentials exist when cohomology vanishes, and interconnection is a gluing operation.
The goal of this note is to make that theme concrete for lumped AC/DC RLC circuits, in a way that is intentionally pedagogical rather than encyclopedic.

\subsection{Charge, conservation, and state variables}
\label{sec:charge-first}
The most robust physical input into circuit theory is conservation of electric charge: currents are rates of charge transfer and Kirchhoff-type constraints are continuity equations.
In the field setting this is the continuity equation
\[
\partial_t\rho + \nabla\cdot \mathbf{j} = 0.
\]
In the lumped graph setting, one can encode the same idea using chains.
Let $X=(V,E)$ be an oriented graph.
Think of vertices as equipotential conductor regions (``nodes'') and let $q(t)\in C_0(X;R)$ denote the (net) charge on each vertex, while $i(t)\in C_1(X;R)$ is the branch current $1$-chain.
Charge conservation can then be written as a discrete continuity equation
\[
\dot q + \partial i = s(t),
\]
where $s(t)$ is an external current-injection $0$-chain supported on terminals.
In regimes where internal charge storage is ignored (or in steady state), $\dot q=0$ at internal nodes and the equation reduces to KCL.

Reactive elements introduce explicit storage coordinates: charge $q$ for capacitors and flux linkage $\Phi$ for inductors.
The universal kinematic relations
\[
i=\frac{dq}{dt},\qquad v=\frac{d\Phi}{dt}
\]
explain why time derivatives appear in constitutive laws, why Fourier characters on $S^1$ produce $i\omega$ factors in steady state, and why Laplace transforms are natural on $[0,\infty)$.
Topology then controls which charge/flux degrees of freedom exist robustly (before geometry supplies $C$ and $L$ as bilinear forms).
% TODO(charge): Tie this charge-first introduction explicitly to Gauss law in the field-domain model and to the Alexander-duality identification of capacitor degrees of freedom.

\subsection{Motivation and a two-layer viewpoint}
Our guiding separation principle is:
\begin{itemize}
  \item \textbf{Topology organizes existence.} (Co)chain complexes, exact sequences, and duality determine \emph{which} degrees of freedom are possible (and their dimensions).
  \item \textbf{Geometry supplies numbers.} A metric and material parameters enter through a Hodge/constitutive operator and energy/dissipation functionals; numerical parameters such as $R$, $L$, and $C$ are matrices of the resulting bilinear forms in physically meaningful bases \cite{jackson_em_1998,bott_tu_1982}.
\end{itemize}

We emphasize two complementary models:
\begin{itemize}
  \item a \emph{lumped graph model} (a finite oriented $1$-complex) capturing Kirchhoff constraints, terminal behavior, and basic compositional rules; and
  \item a \emph{field-domain model} $U=S^3\setminus \mathrm{int}(M)$ capturing the topological degrees of freedom behind capacitance and inductance via Alexander duality and Poincar\'e--Lefschetz duality.
\end{itemize}

\subsection{Related work (brief)}
This project is strongly influenced by the compositional/variational viewpoint of Baez--Fong and collaborators on passive linear networks \cite{baez_fong_2015_compositional,baez_fong_pollard_2016}.
Our emphasis differs mainly in that we keep the ambient $3$-dimensional topology in view (to explain energy storage and linking phenomena) and we treat time periodicity via a space--time product picture (rather than, say, Laplace transform as a primary language).

For topological electromagnetics and the use of duality/linking in computational settings, see Gross--Kotiuga \cite{gross_kotiuga_2004}.
For the functional-analytic/discretization perspective where Hodge theory, boundary conditions, and spectral gaps matter (useful for the ``almost topological'' discussion), see finite element exterior calculus (FEEC) \cite{arnold_falk_winther_2010_feec}.

\subsection{Organization and reading guide}
Section~\ref{sec:background} collects notation and the specific algebraic-topology tools we use.
Section~\ref{sec:lumped-graph} develops the lumped graph model and proves the Kirchhoff statements.
Section~\ref{sec:field-domain} introduces the field-domain model, terminals, and measurement as duality pairings.
Section~\ref{sec:gluing} records excision and Mayer--Vietoris and explains how Dirichlet-to-Neumann maps and Schur complements encode ``black-box'' behavior under interconnection.
Sections~\ref{sec:alexander-duality} and~\ref{sec:time-topology} treat the two main organizing mechanisms for RLC behavior beyond Kirchhoff bookkeeping: Alexander duality (capacitive/inductive state spaces) and time topology (AC vs.\ DC).
Section~\ref{sec:examples} contains worked examples and an optional ``almost topological'' module.

\subsection{Main statements (informal but precise)}
\label{sec:main-statements}
The following statements are the spine of the paper; later sections provide proof sketches and examples.

\begin{theorem}[Kirchhoff laws as boundary/coboundary conditions]\label{thm:kcl-kvl}
Let $X=(V,E)$ be a finite oriented graph and let $T\subset V$ be a set of \emph{terminal} vertices.
Fix a coefficient field $R$ (typically $R=\mathbb{R}$).
Currents are $1$-chains $i\in C_1(X;R)$ and voltages are $1$-cochains $v\in C^1(X;R)$.
Then:
\begin{itemize}
  \item \textbf{(KCL)} Current conservation at internal nodes is the condition
  \[
  \partial i = 0 \quad \text{in } C_0(X,T;R),
  \]
  i.e.\ $\partial i$ is supported on terminals.
  \item \textbf{(KVL)} The existence of a node potential $\phi\in C^0(X;R)$ with $v=d\phi$ is equivalent to
  \[
  v(\gamma)=0 \quad \text{for all cycles } \gamma\in Z_1(X;R),
  \]
  i.e.\ to $[v]=0$ in $H^1(X;R)$.
\end{itemize}
\end{theorem}

\begin{remark}[Charge conservation behind KCL]\label{rem:charge-kcl}
In time-dependent settings, KCL is a special case of charge conservation (Section~\ref{sec:charge-first}).
Writing vertex charges as a time-dependent $0$-chain $q(t)\in C_0(X;R)$ and currents as $i(t)\in C_1(X;R)$, the discrete continuity equation is
\[
\dot q + \partial i = s(t)
\]
with $s(t)$ supported on terminals (external injection).
When internal charge accumulation is ignored, $\dot q=0$ on internal vertices, so $\partial i=0$ in $C_0(X,T;R)$, which is exactly KCL as stated above.
% TODO(charge): Make the capacitor case explicit here by expressing $q$ in terms of terminal/node potentials via the capacitance operator and showing $i=dq/dt$.
\end{remark}

\begin{theorem}[AC adds time-circle degrees of freedom]\label{thm:ac-kunneth}
Let $Y$ be a space (or pair) modeling spatial degrees of freedom (e.g.\ $(X,T)$ or $(U,\Gamma)$), and let $\mathcal{T}$ be either the half-line $[0,\infty)$ (or its one-point compactification $[0,\infty]$ for topology) or a circle $S^1$ (steady periodic forcing).
Over a field, the (co)homology of $Y\times \mathcal{T}$ decomposes via K\"unneth as a direct sum of tensor products.
In particular, replacing a contractible time axis by a time circle adds a new $H_1(S^1)$ factor, which is the topological shadow of additional AC/phasor degrees of freedom.
\end{theorem}

\begin{theorem}[Capacitor/inductor degrees of freedom via Alexander duality]\label{thm:alexander-rlc}
Let $M\subset S^3$ be a compact conductor region with field domain $U=S^3\setminus \mathrm{int}(M)$.
Then the degrees of freedom governing electrostatics and magnetostatics are naturally expressed in terms of reduced homology of $M$ and (co)homology of $U$, related by Alexander duality.
With terminals $\Gamma\subset \partial U$ specified, the relevant spaces are relative groups and the associated Poincar\'e--Lefschetz pairings.
\end{theorem}

\begin{theorem}[Measurement as a duality pairing]\label{thm:measurement-pairing}
Let $U$ be an oriented compact $3$-manifold-with-boundary and let $\Gamma\subset \partial U$ be a union of terminal boundary components.
Poincar\'e--Lefschetz duality supplies perfect evaluation pairings of the form
\[
H^k(U,\Gamma;R)\times H_{3-k}(U,\partial U\setminus \Gamma;R)\longrightarrow R,
\]
which can be interpreted as ``effort evaluated on flow''.
After choosing a constitutive Hodge operator $\star_{\mathrm{mat}}$, these pairings induce inner products and hence canonical projections/minimization principles underlying idealized readout.
\end{theorem}

\begin{theorem}[Gluing, exact sequences, and Schur complements]\label{thm:gluing-schur}
Suppose a device is built by gluing two subdevices along an interface (at the level of graphs or field domains).
Then Mayer--Vietoris and related long exact sequences describe how the relevant (co)homology state spaces split or are constrained under gluing.
After choosing a quadratic energy/dissipation functional on those topologically determined state spaces, eliminating internal interface degrees of freedom produces a Schur complement, yielding an effective terminal operator (a Dirichlet-to-Neumann map) compatible with interconnection.
\end{theorem}

\section{Background, notation, and conventions}
\label{sec:background}
We assume basic familiarity with singular/cellular (co)homology, long exact sequences, and duality; standard references include \cite{hatcher_at_2002,munkres_topology_2000}.
For differential forms and the Hodge-star viewpoint used to connect topology to energy pairings, see \cite{bott_tu_1982,morita_geometry_of_diff_forms_2001}.

\subsection{Notation and coefficients}
Unless stated otherwise we work over $R=\mathbb{R}$.
In AC steady state we sometimes extend scalars to $\mathbb{C}$ (phasors).
We write $\widetilde H_\bullet$ and $\widetilde H^\bullet$ for reduced (co)homology.
We reserve $T\subset V$ for graph terminals, and use $\mathcal{T}$ for the time manifold factor (typically $S^1$ or $[0,\infty)$; sometimes its one-point compactification $[0,\infty]$ is used for compact-topology arguments).

\subsection{Chains and cochains on a graph}
Throughout, a lumped circuit is modeled as a finite oriented $1$-dimensional CW complex (a graph)
\[
X=(V,E).
\]
Cellular chain groups are $C_k(X;R)$ with boundary $\partial\colon C_1(X;R)\to C_0(X;R)$, and cochains are $C^k(X;R)=\mathrm{Hom}(C_k(X;R),R)$ with coboundary $d\colon C^0(X;R)\to C^1(X;R)$.
Evaluation provides the basic pairing $C^1(X;R)\times C_1(X;R)\to R$.

\paragraph{Charge and the discrete continuity equation (optional).}
If one includes capacitive storage explicitly, it is convenient to introduce a vertex-charge variable $q(t)\in C_0(X;R)$ alongside the branch-current $1$-chain $i(t)\in C_1(X;R)$.
Charge conservation is then the continuity equation $\dot q+\partial i=s(t)$ with $s(t)$ supported on terminals.
See Section~\ref{sec:charge-first} for the ``charge-first'' viewpoint.

\subsection{Relative (co)homology and terminals}
\paragraph{Graph terminals.}
To speak about a \emph{device} rather than a closed circuit, we fix a set of terminal vertices $T\subset V$, viewed as a subcomplex.
Relative chain groups $C_\bullet(X,T;R)$ encode ``internal conservation'' with boundary allowed on terminals; for example KCL at internal nodes is $\partial i=0$ in $C_0(X,T;R)$.
The relative groups $H_\bullet(X,T;R)$ and $H^\bullet(X,T;R)$ carry a natural Kronecker pairing
\[
H^k(X,T;R)\times H_k(X,T;R)\to R,
\]
which is the clean home for ``readouts between terminals''.

\paragraph{Field terminals.}
In the field-domain model, terminals are represented by a union $\Gamma\subset \partial U$ of boundary components where boundary conditions/probes are imposed.
The pair $(U,\Gamma)$ determines relative groups such as $H^\bullet(U,\Gamma;R)$ and $H_\bullet(U,\partial U\setminus \Gamma;R)$, which are dual under Poincar\'e--Lefschetz duality.

\subsection{Exact-sequence toolkit}
We use:
\begin{itemize}
  \item the long exact sequence of a pair $(X,A)$,
  \item excision (local modifications away from terminals do not change relative groups), and
  \item Mayer--Vietoris long exact sequences for decompositions $X=A\cup B$,
\end{itemize}
to formalize how constraints and degrees of freedom behave under interconnection.
We also use functoriality: inclusions induce maps on (co)homology, and most constructions below are natural with respect to these maps.

\subsection{Dualities used in the paper}
\paragraph{Alexander duality (in $S^3$).}
For a compact locally contractible $A\subset S^3$ and coefficient ring/field $R$, Alexander duality provides
\[
\widetilde H^k(S^3\setminus A;R)\cong \widetilde H_{2-k}(A;R).
\]

\paragraph{Poincar\'e--Lefschetz duality.}
For an oriented compact $3$-manifold-with-boundary $U$ and $\Gamma\subset \partial U$, there are perfect pairings
\[
H^k(U,\Gamma;R)\times H_{3-k}(U,\partial U\setminus \Gamma;R)\to R
\]
given by evaluation/integration.

\subsection{Cup, cap, and slant products (optional)}
The most frequently used ``product'' in this note is evaluation: a cohomology class is measured on a homology class.
At the level of relative groups this is the Kronecker pairing
\[
H^k(X,A;R)\times H_k(X,A;R)\to R,
\]
and more generally it can be expressed using the cap product (and, dually, cup products) \cite{hatcher_at_2002,bott_tu_1982}.

\paragraph{Cup products (where they can matter).}
On a graph $X$ (a $1$-complex), $H^2(X;R)=0$, so the cup product of two $H^1$ classes necessarily vanishes.
In contrast, for a $3$-dimensional field domain $U$ one can have nontrivial products such as
\[
H^1(U;R)\times H^1(U;R)\to H^2(U;R),
\]
and under Poincar\'e--Lefschetz duality these products are closely related to intersection/linking-type pairings (heuristically: ``multiply two circulations to get a flux class'').

\paragraph{Slant products (integration along a factor).}
For a product $Y\times S^1$, K\"unneth describes how classes decompose into spatial and temporal factors.
The slant product with the fundamental class $[S^1]\in H_1(S^1;R)$ provides a canonical ``integration over time'' map
\[
H^{k+1}(Y\times S^1;R)\to H^{k}(Y;R),
\]
which is the algebraic-topology avatar of taking a period integral / time-averaged quantity in the de Rham picture.
We return to this viewpoint in the AC discussion in \S\ref{sec:dc-vs-ac-topology}.

\subsection{Hodge star and constitutive data}
Topology alone produces (co)homology groups describing \emph{which} fluxes/charges/circulations are possible.
Numerical parameters require geometric/constitutive data, typically encoded by a Hodge/constitutive operator $\star_{\mathrm{mat}}$ and the associated inner product
\[
\langle \alpha,\beta\rangle_{\mathrm{mat}}:=\int_U \alpha\wedge (\star_{\mathrm{mat}}\beta).
\]

\subsection{Discrete Hodge theory on cell complexes (optional)}
\label{sec:discrete-hodge}
The same Hodge-theoretic package used in the smooth setting has a finite-dimensional analogue on a cell complex.
This is a clean way to state that ``solving a linear circuit is a harmonic boundary-value problem plus topological zero-modes.''

\paragraph{Cochains, adjoints, and the Hodge Laplacian.}
Let $K$ be a finite cell complex and consider the cochain complex $(C^\bullet(K;R),d)$.
Choose inner products $\langle\cdot,\cdot\rangle_k$ on each $C^k(K;R)$ (in circuits these are induced by element parameters and play the role of a discrete constitutive/Hodge operator).
Let $\delta\colon C^{k}(K;R)\to C^{k-1}(K;R)$ denote the adjoint of $d$ with respect to these inner products.
The (combinatorial) Hodge Laplacian is
\[
\Delta_k := d\,\delta + \delta\,d\qquad\text{on }C^{k}(K;R).
\]
It is self-adjoint and positive semidefinite.

\paragraph{Harmonic cochains and cohomology.}
Define the harmonic $k$-cochains by $\mathcal{H}^k(K):=\ker \Delta_k$.
In this finite-dimensional setting one has a discrete Hodge theorem: harmonic cochains represent cohomology classes, and (after choosing boundary conditions) there are natural isomorphisms
\[
\mathcal{H}^k(K)\;\cong\; H^k(K;R),
\]
and similarly for relative cohomology when one imposes terminal/boundary conditions (Dirichlet/Neumann analogues).
See, e.g., the FEEC viewpoint \cite{arnold_falk_winther_2010_feec}.

\paragraph{Hodge decomposition and a spectral basis.}
One also has the orthogonal Hodge decomposition
\[
C^k(K;R) \;=\; \image(d)\;\oplus\;\mathcal{H}^k(K)\;\oplus\;\image(\delta).
\]
Equivalently, every $k$-cochain splits uniquely into \emph{exact} ($d$-part), \emph{harmonic} (topological), and \emph{coexact} ($\delta$-part) components.
Because $\Delta_k$ preserves $\image(d)$ and $\image(\delta)$ (it commutes with $d$ and $\delta$), one can choose an orthonormal eigenbasis of $\Delta_k$ adapted to this decomposition:
the $\lambda=0$ eigenvectors are precisely the harmonic cochains, and the $\lambda>0$ eigenvectors span the exact/coexact subspaces (with the chosen boundary conditions).

\paragraph{Circuit interpretation (what corresponds to KCL/KVL).}
In the lumped setting the relevant complex is often a $1$-complex $X$ (a graph) or a low-dimensional cell complex built from it.
Under the usual identifications, the operator $\delta$ is a weighted ``divergence'' and $d$ is a discrete ``gradient''.
Then:
\begin{itemize}
  \item \emph{Electrostatic KVL:} a voltage $1$-cochain is exact, $v=d\phi$ (no harmonic circulation component).
  \item \emph{KCL (internal nodes):} a current is co-closed, $\delta i=0$, i.e.\ divergence-free on internal nodes (with terminals treated by relative/boundary conditions).
\end{itemize}
The harmonic part is the topology-controlled remainder: it is the space of zero-modes that survives all local constraints, and its dimension is a Betti number (absolute or relative, depending on terminals).

\subsection{\texorpdfstring{Time topology: $[0,\infty]$ versus $S^1$}{Time topology: interval versus circle}}
We model time dependence by products with a $1$-manifold $\mathcal{T}$:
\begin{itemize}
  \item \emph{DC/transients:} $\mathcal{T}=[0,\infty)$ (or compactified to $[0,\infty]$ when using compact-topology arguments), and
  \item \emph{AC steady periodic forcing:} $\mathcal{T}=S^1$ (periodic, no boundary).
\end{itemize}
K\"unneth decompositions describe how state spaces change under $Y\mapsto Y\times \mathcal{T}$.
A useful caution is that homology of $S^1$ only sees the constant class and the winding class; the full Fourier spectrum of a periodic signal enters analytically (as eigenfunctions/characters), not through singular homology alone.

\section{The lumped graph model}
\label{sec:lumped-graph}
\subsection{Model and dictionary}
Let $X=(V,E)$ be a finite oriented graph.
We interpret:
\begin{itemize}
  \item branch currents as $1$-chains $i\in C_1(X;R)$,
  \item node potentials as $0$-cochains $\phi\in C^0(X;R)$, and
  \item branch voltages as $1$-cochains $v\in C^1(X;R)$.
\end{itemize}
If terminals $T\subset V$ are specified, then $C_\bullet(X,T;R)$ encodes ``internal'' conservation with boundary allowed on terminals.

\begin{center}
\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
\hline
Circuit notion & Topological/algebraic notion \\
\hline
Branch current $i$ & $1$-chain $i\in C_1(X;R)$ \\
Node current injection & $\partial i\in C_0(X;R)$ (supported on $T$) \\
Node potential $\phi$ & $0$-cochain $\phi\in C^0(X;R)$ \\
Branch voltage $v$ & $1$-cochain $v\in C^1(X;R)$ \\
Voltage from a potential & $v=d\phi$ \\
KCL (internal) & $\partial i=0$ in $C_0(X,T;R)$ \\
KVL (no induction) & $v$ is exact; $[v]=0$ in $H^1(X;R)$ \\
\hline
\end{tabular}
\end{center}

\subsection{Kirchhoff laws and potentials}
\paragraph{Proof sketch of \Cref{thm:kcl-kvl}.}
The chain complex of a graph is $0\to C_1(X;R)\xrightarrow{\partial} C_0(X;R)\to 0$.
Thus $\partial i=0$ is exactly the statement that currents are $1$-cycles (and $\partial i=0$ in $C_0(X,T;R)$ is the relative version allowing sources at terminals).
For voltages, the coboundary $d\colon C^0(X;R)\to C^1(X;R)$ is the transpose incidence map in a chosen basis.
If $v=d\phi$, then $v(\gamma)=0$ on every cycle $\gamma$ (telescoping sum).
Conversely, if $v$ evaluates to $0$ on all cycles, then integrating $v$ along a path from a base vertex defines $\phi$; the cycle condition ensures path independence, so $v=d\phi$.

\paragraph{Discrete Hodge viewpoint (optional).}
With a choice of inner products on cochains (coming from element parameters), this same picture can be rephrased using the discrete Hodge Laplacian and Hodge decomposition: KVL selects the exact part of a voltage $1$-cochain, KCL is co-closedness (a divergence-free condition) for currents, and the remaining harmonic subspace is canonically isomorphic to cohomology \cite{arnold_falk_winther_2010_feec}.
See \S\ref{sec:discrete-hodge}.

\subsection{Terminals and ``voltage between terminals'' as a relative pairing}
Fix terminals $T\subset V$.
A probe path between two terminals is modeled by a $1$-chain $\gamma$ with $\partial\gamma\in C_0(T;R)$.
The voltage readout along that probe is $v(\gamma)$.
If $v=d\phi$ then $v(\gamma)=\phi(t_1)-\phi(t_0)$ depends only on the terminal endpoints.
More generally, relative homology $H_1(X,T;R)$ packages paths modulo deformations through the interior, and the Kronecker pairing with $H^1(X,T;R)$ is the natural topological home for terminal readouts.

\subsection{Worked example: a one-loop graph (explicit matrices)}
\label{sec:example-tree-loop}
Consider the ``parallel-edge'' graph with two vertices $v_0,v_1$ and two edges $e_1,e_2$ oriented from $v_0$ to $v_1$:
\begin{center}
\begin{tikzpicture}[
  scale=1.5,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  \tikzset{
    vnode/.style={circle,draw,inner sep=2.6pt,line width=0.85pt,fill=white},
    epath/.style={->,line width=1.1pt,shorten <=1.6pt,shorten >=1.6pt},
    elab/.style={fill=white,fill opacity=0.85,text opacity=1,inner sep=1.2pt,rounded corners=1pt}
  }
  \node[vnode] (v0) at (0,0) {$v_0$};
  \node[vnode] (v1) at (3.2,0) {$v_1$};
  \draw[epath] (v0) to[out=20,in=160] node[pos=0.55,elab,above] {$e_1$} (v1);
  \draw[epath] (v0) to[out=-20,in=-160] node[pos=0.55,elab,below] {$e_2$} (v1);
\end{tikzpicture}
\end{center}
\paragraph{KCL (cycle space).}
With bases $(e_1,e_2)$ for $C_1\cong R^2$ and $(v_0,v_1)$ for $C_0\cong R^2$, the boundary map is
\[
\partial=\begin{pmatrix}-1 & -1\\ 1 & 1\end{pmatrix},
\]
so for a current $i=(i_1,i_2)\in C_1(X;R)$ we have
\[
\partial(i_1,i_2)=(-i_1-i_2)\,v_0+(i_1+i_2)\,v_1.
\]
If there are no external injections at the nodes (equivalently, $T=\emptyset$ and KCL is $\partial i=0$), then $i_2=-i_1$.
Thus the cycle space is
\[
Z_1(X;R)=\ker\partial=\mathrm{span}\{(1,-1)\},
\]
which is rank~$1$ (one independent loop current).

\paragraph{KVL (exact voltages).}
The coboundary $d\colon C^0\to C^1$ is represented by the transpose matrix
\[
d=\partial^\top=\begin{pmatrix}-1 & 1\\ -1 & 1\end{pmatrix},
\]
so $\mathrm{im}(d)=\mathrm{span}\{(1,1)\}\subset C^1\cong R^2$.
Concretely, writing $\phi=(\phi_0,\phi_1)\in C^0(X;R)$ gives
\[
d\phi=(\phi_1-\phi_0,\;\phi_1-\phi_0),
\]
so a voltage assignment $v=(v_1,v_2)$ is exact (hence satisfies electrostatic KVL) if and only if $v_1=v_2$.

\paragraph{A cohomology obstruction as ``induced EMF''.}
Since $H^1(X;R)\cong C^1/\mathrm{im}(d)\cong R$, there are non-exact $1$-cochains.
For instance, $v=(1,-1)$ is not exact.
Evaluating on the fundamental cycle $\gamma=(1,-1)\in Z_1$ gives the loop integral
\[
v(\gamma)= v_1\gamma_1+v_2\gamma_2 = 1\cdot 1 + (-1)\cdot (-1)=2\neq 0.
\]
Engineering interpretation: a nonzero loop integral represents an electromotive force around the loop (e.g.\ due to time-varying magnetic flux), i.e.\ precisely the situation in which the electrostatic KVL assumption fails.

\section{Field-domain model and measurement}
\label{sec:field-domain}
This section introduces the ambient $3$-dimensional setup needed to explain energy storage and idealized measurement in genuinely topological terms.

\subsection{Conductor regions and the field domain}
Let $S^3$ denote the one-point compactification of $\mathbb{R}^3$.

\begin{definition}[Conductor region]
A \emph{(perfect) conductor region} is a compact $3$-dimensional submanifold-with-boundary
\[
M\subset S^3
\]
whose connected components are the individual conducting bodies (wires, plates, coils, etc.).
We write
\[
\partial M=\bigsqcup_{\alpha=1}^m \Sigma_\alpha
\]
for its boundary components.
\end{definition}

\begin{definition}[Field domain]
The \emph{field domain} (exterior region) is
\[
U:=S^3\setminus \mathrm{int}(M),
\]
which is a compact $3$-manifold-with-boundary with $\partial U=\partial M$.
\end{definition}

\paragraph{Charge as data on conductor components.}
In electrostatics, the total charge on each connected conductor component is a primary state/readout.
At the ideal level one may package these charges as a vector
\[
Q=(Q_1,\dots,Q_m)\in R^m\cong H_0(M;R)^\ast,
\]
where $Q_\alpha$ is the net charge on the component corresponding to $\Sigma_\alpha$ (equivalently, on the associated component of $M$).
Because only charge differences are physically meaningful in an isolated device (global offsets correspond to an overall ``background'' charge), the reduced group $\widetilde H_0(M;R)$ is the natural home for capacitor-type storage coordinates.
% TODO(charge): Make the Gauss-law link explicit: relate $Q$ to electric flux through separating surfaces in $U$ and to the Alexander-duality identification $\widetilde H_0(M)\cong \widetilde H^2(U)$.
% TODO(charge): Connect this to the DtN/capacitance-matrix construction: terminal potentials determine induced charges by energy minimization.

\begin{remark}
The $S^3$-model avoids boundary conditions at spatial infinity; one can instead work in $\mathbb{R}^3$ with decay conditions without changing the purely topological statements below.
\end{remark}

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  scale=1.0,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  % Ambient cross-section (schematic $S^3$)
  \shade[ball color=blue!10] (0,0) circle (2.2);
  \draw[line width=0.95pt] (0,0) circle (2.2);
  \node at (0,2.58) {$S^3$ (schematic)};

  % Conductor components (filled) and their boundaries (which are $\partial U=\partial M$)
  \shade[ball color=gray!28] (-0.95,0.45) circle (0.60);
  \shade[ball color=gray!28] (0.95,-0.25) circle (0.60);
  \draw[line width=0.95pt] (-0.95,0.45) circle (0.60);
  \draw[line width=0.95pt] (0.95,-0.25) circle (0.60);
  \node at (-0.95,0.45) {$M_1$};
  \node at (0.95,-0.25) {$M_2$};

  \node[
    blue!70!black,
    fill=white,
    fill opacity=0.75,
    text opacity=1,
    rounded corners=2pt,
    inner sep=2pt
  ] at (0,-1.62) {$U=S^3\setminus\mathrm{int}(M)$};
  \draw[->,gray!70!black,line width=0.9pt] (2.05,0.95) -- (1.45,0.12);
  \node[
    gray!70!black,
    fill=white,
    fill opacity=0.75,
    text opacity=1,
    rounded corners=2pt,
    inner sep=2pt
  ] at (2.22,1.25) {$\partial U=\partial M$};

  % Highlight a terminal subset $\Gamma\subset\partial U$
  \draw[line width=7pt,red!25,opacity=0.55] (-0.95,0.45) ++(35:0.60) arc (35:155:0.60);
  \draw[line width=1.6pt,red!80!black] (-0.95,0.45) ++(35:0.60) arc (35:155:0.60);
  \draw[->,red!80!black,line width=0.9pt] (-1.78,1.52) -- (-1.25,1.05);
  \node[
    red!80!black,
    fill=white,
    fill opacity=0.8,
    text opacity=1,
    rounded corners=2pt,
    inner sep=1.6pt
  ] at (-2.05,1.63) {$\Gamma$};
\end{tikzpicture}
\caption{A schematic of the conductor region $M\subset S^3$ and the field domain $U=S^3\setminus\mathrm{int}(M)$.
Terminals are modeled as a distinguished subset $\Gamma\subset \partial U=\partial M$ where boundary conditions and probes are imposed.}
\label{fig:field-domain-setup}
\end{figure}

\subsection{Terminals and relative groups}
To model a component, one must specify which boundary components are treated as terminals.
Accordingly, we allow a decomposition
\[
\partial U = \Gamma \sqcup (\partial U\setminus \Gamma),
\]
where $\Gamma$ is the union of terminal surfaces.
The relevant state and measurement spaces are relative groups $H^\bullet(U,\Gamma;R)$ and $H_\bullet(U,\partial U\setminus \Gamma;R)$, paired by Poincar\'e--Lefschetz duality.

\subsection{Measurement as constraint + pairing}
\paragraph{Topological content (pairing).}
Idealized measurement has two parts: a constraint (``do not disturb'') and a readout.
The readout is topological: it is evaluation of an ``effort'' class on a ``flow'' class, as in \Cref{thm:measurement-pairing}.

\paragraph{Graph level (lumped probes).}
In the graph model, a voltmeter between terminals can be represented by a relative $1$-cycle class in $H_1(X,T)$ (a probe path), and its readout is evaluation of a voltage $1$-cochain on that class.
An ammeter is represented dually by choosing a cut/edge and evaluating a current $1$-chain.

\paragraph{Field level (Poincar\'e--Lefschetz + Hodge).}
In the field model, a voltmeter corresponds to evaluating a (relative) electric $1$-form class $[E]\in H^1(U,\Gamma)$ on a relative $1$-cycle $\gamma\in H_1(U,\partial U\setminus \Gamma)$:
\[
\mathrm{Volt}([E],\gamma)=\int_\gamma E.
\]
An ammeter corresponds to evaluating a current-flux $2$-class on an appropriate spanning surface.
Geometry/material data enters through $\star_{\mathrm{mat}}$, which turns these evaluation pairings into inner products and variational principles (minimization of energy or dissipation).

\subsection{Poincar\'e--Lefschetz pairing vs.\ Hodge star (why measurement needs a metric)}
Poincar\'e--Lefschetz duality provides a \emph{topological} pairing between effort-type and flow-type state spaces:
\[
H^k(U,\Gamma;R)\times H_{3-k}(U,\partial U\setminus \Gamma;R)\to R,
\]
defined by evaluation/integration on representatives.
This pairing is metric-free: it exists as soon as the relevant (co)homology groups are defined.

The Hodge star enters at a different level.
Once a metric and material parameters are chosen, the constitutive Hodge operator $\star_{\mathrm{mat}}$ identifies $k$-forms with $(3-k)$-forms and turns evaluation into an inner product:
\[
\langle \alpha,\beta\rangle_{\mathrm{mat}}:=\int_U \alpha\wedge(\star_{\mathrm{mat}}\beta).
\]
In Hodge theory, harmonic representatives identify cohomology with kernels of Laplacians, and $\star_{\mathrm{mat}}$ maps harmonic $k$-forms to harmonic $(3-k)$-forms; in this sense it realizes Poincar\'e--Lefschetz duality as a concrete ``Hodge dual'' \cite{bott_tu_1982,arnold_falk_winther_2010_feec}.
But unlike Poincar\'e--Lefschetz duality, $\star_{\mathrm{mat}}$ is not topological: it depends on geometry and material calibration, which is why it is the right home for numerical readout and ``least disturbance'' measurement principles.

\paragraph{Periods and gauge/constraints.}
If $\alpha$ is a closed $k$-form and $\gamma$ is a $k$-cycle, then the \emph{period}
\[
\int_\gamma \alpha
\]
depends only on the cohomology class $[\alpha]\in H^k$ and the homology class $[\gamma]\in H_k$.
Periods are therefore the canonical invariants extracted by measurement pairings.

In circuit modeling one often sets certain periods to zero by assumption.
For example, electrostatic KVL corresponds to restricting to exact electric fields (vanishing periods around loops), i.e.\ ``no induced EMF''.
Similarly, overall potential offsets (the reduced $H^0$ direction) are fixed by grounding conventions.
More generally, ``set to zero'' enters through boundary conditions at terminals, gauge fixing, or by restricting attention to a subspace of classes appropriate to the regime (electrostatic vs.\ magnetostatic vs.\ fully time-dependent).

\subsection{From topology to numbers: Hodge and energy}
\label{sec:elements-as-pairings}
The topology of $U$ determines finite-dimensional state spaces (ranks of the relevant (co)homology groups).
Numerical parameters arise after choosing material data, schematically represented by functions/tensors
\[
\epsilon\colon U\to \mathbb{R}_{>0},\qquad \mu\colon U\to \mathbb{R}_{>0},\qquad \sigma\colon U\to \mathbb{R}_{\ge 0}.
\]
These are \emph{not} coefficients of singular cohomology; rather they build bilinear forms by integration.
For example, the material Hodge operator $\star_{\mathrm{mat}}$ yields a symmetric form
\[
([\alpha],[\beta])\longmapsto \int_U \alpha\wedge(\star_{\mathrm{mat}}\beta)
\]
on the appropriate cohomology group.
Choosing bases of topological classes (charges/fluxes/loops) turns these forms into matrices; those matrices are the lumped $C$ and $L$ parameters.
Similarly, dissipation $\int_U \sigma |E|^2\,dV$ produces effective resistive operators.

\section{Gluing and effective elements}
\label{sec:gluing}
Circuit building is a gluing operation: we form a larger device by joining subdevices along interfaces.
On the topological side, the bookkeeping tools are excision and Mayer--Vietoris.
On the analytic side, effective ``black-box'' behavior is encoded by Dirichlet-to-Neumann maps and Schur complements.
The intended workflow in this section is topology-first: first identify the global state space and compatibility constraints by exact-sequence methods, and only then insert constitutive quadratic functionals to obtain numerical reduction formulas.

\subsection{Excision (local modifications do not change relative groups)}
\label{sec:excision}
Excision says that relative (co)homology is insensitive to removing subsets away from the region where boundary conditions are imposed.
Informally, if $(X,A)$ is a pair and $B\subset A$ lies in the interior of $A$, then the inclusion
\[
(X\setminus B,\; A\setminus B)\hookrightarrow (X,A)
\]
induces isomorphisms on relative homology and cohomology.
In our setting, this justifies replacing complicated conductor shapes by deformation retracts and simplifying $U$ away from terminals, as long as we keep fixed the terminal region $\Gamma$.

\subsection{Mayer--Vietoris (cut-and-paste exact sequences)}
\label{sec:mayervietoris}
Let $X=A\cup B$ be a union of subspaces with ``nice'' intersection $A\cap B$.
Then Mayer--Vietoris provides a long exact sequence relating $H_\bullet(X)$ to $H_\bullet(A)$, $H_\bullet(B)$, and $H_\bullet(A\cap B)$ (and similarly for cohomology).

Concretely, for homology one has an exact sequence of the form
\[
\cdots \longrightarrow H_n(A\cap B)\longrightarrow H_n(A)\oplus H_n(B)\longrightarrow H_n(X)\longrightarrow H_{n-1}(A\cap B)\longrightarrow \cdots,
\]
and for cohomology one has
\[
\cdots \longrightarrow H^{n-1}(A\cap B)\longrightarrow H^{n}(X)\longrightarrow H^{n}(A)\oplus H^{n}(B)\longrightarrow H^{n}(A\cap B)\longrightarrow \cdots.
\]

\paragraph{Interpretation for circuits.}
If a device is decomposed into subdevices whose field domains overlap on an interface, Mayer--Vietoris encodes compatibility constraints and identifies how independent degrees of freedom in each piece combine (or are killed) in the union.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  scale=1.0,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  % Two overlapping pieces with a highlighted overlap
  \def\R{1.55}
  \def\d{2.0}
  \pgfmathsetmacro{\theta}{acos((\d/2)/\R)} % degrees
  \coordinate (Ac) at (-1.0,0);
  \coordinate (Bc) at (1.0,0);

  \shade[inner color=blue!6,outer color=blue!20,opacity=0.55] (Ac) circle (\R);
  \shade[inner color=green!8,outer color=green!22,opacity=0.45] (Bc) circle (\R);
  \begin{scope}
    \clip (Ac) circle (\R);
    \fill[purple!70,opacity=0.12] (Bc) circle (\R);
  \end{scope}

  \draw[line width=1.0pt,blue!70!black] (Ac) circle (\R);
  \draw[line width=1.0pt,green!60!black] (Bc) circle (\R);

  % Lens boundary (the interface)
  \draw[line width=1.1pt,purple!75!black]
    (Ac) ++({\theta}:\R) arc ({\theta}:{-\theta}:\R);
  \draw[line width=1.1pt,purple!75!black]
    (Bc) ++({180-\theta}:\R) arc ({180-\theta}:{180+\theta}:\R);

  \node[blue!70!black] at (-2.05,1.35) {$A$};
  \node[green!60!black] at (2.05,1.35) {$B$};

  \node[
    draw=purple!75!black,
    rounded corners=2pt,
    fill=white,
    fill opacity=0.88,
    text opacity=1,
    inner sep=2pt,
    align=center
  ] at (0,0) {$A\cap B$\\[-1pt]\scriptsize interface};
  \node[fill=white,fill opacity=0.75,text opacity=1,rounded corners=2pt,inner sep=1.6pt] at (0,-2.15) {$X=A\cup B$};
\end{tikzpicture}
\caption{A Mayer--Vietoris decomposition $X=A\cup B$ with overlap $A\cap B$ representing an interface for gluing.}
\label{fig:mv-decomposition}
\end{figure}

\begin{proposition}[Direct sums vs.\ constraints (a Mayer--Vietoris dichotomy)]\label{prop:mv-dichotomy}
In favorable decompositions, the relevant part of the Mayer--Vietoris sequence reduces to one of two patterns:
\begin{enumerate}
  \item \emph{Direct sum.} If the overlap $A\cap B$ is homologically trivial in the degree of interest, then $H_k(X)$ (or $H^k(X)$) is isomorphic to a direct sum of the corresponding groups of $A$ and $B$.
  \item \emph{Matching constraint.} If the overlap carries nontrivial data in the degree of interest, then $H^k(X)$ is naturally identified with a kernel (or quotient) enforcing a matching condition across the interface, e.g.\ as a kernel of a ``restriction-difference'' map $H^k(A)\oplus H^k(B)\to H^k(A\cap B)$.
\end{enumerate}
\end{proposition}

\noindent\emph{Proof sketch.}
Both patterns are immediate from exactness: if $H_\bullet(A\cap B)$ vanishes in the relevant degrees, then the connecting maps in Mayer--Vietoris force the middle map to be an isomorphism (direct sum).
If instead the overlap is nontrivial, exactness identifies the global group with the appropriate kernel/quotient enforcing consistency of restrictions on $A\cap B$.

\subsection{A topology-first reduction workflow}
\label{sec:topology-first-workflow}
The point of the gluing formalism is not merely to justify a later Schur-complement computation, but to separate \emph{structural} questions from \emph{numerical} ones.
In practice one can organize reductions in three steps:
\begin{enumerate}
  \item \emph{Excision/simplification:} replace complicated geometry away from terminals by a deformation retract or another excisively equivalent model.
  \item \emph{Exact-sequence bookkeeping:} use Mayer--Vietoris (and pair exact sequences) to determine whether state spaces glue by direct sum, kernel, or quotient and to identify the terminal/interface constraint maps.
  \item \emph{Constitutive reduction:} choose a Hodge/material inner product or quadratic energy/dissipation functional on the resulting finite-dimensional state spaces and eliminate internal variables (DtN/Schur complement).
\end{enumerate}
Steps (1)--(2) are purely topological and determine \emph{which} reduction laws are even meaningful; step (3) supplies the coefficients and frequency scales.

\subsection{Dirichlet-to-Neumann maps and Schur complements}
\label{sec:dtn}
\paragraph{Proof idea for \Cref{thm:gluing-schur}.}
After the gluing step has identified the relevant interface/terminal variables topologically, a quadratic functional in the bulk (energy or dissipation) produces an induced quadratic form on the terminal space by fixing terminal data and minimizing in the interior.
This is a Dirichlet-to-Neumann (DtN) construction; gluing corresponds to eliminating interface variables, which is a Schur complement.

\paragraph{Resistive DtN (effective conductance/resistance).}
For dissipation $\mathcal{D}(E)=\int_U \sigma |E|^2\,dV$, minimizing subject to terminal potentials yields a map ``terminal voltage $\mapsto$ terminal current''; effective resistance is the inverse of this map on the relevant subspace \cite{jackson_em_1998,chua_desoer_kuh_1987}.

\paragraph{Capacitive and inductive DtN (capacitance/inductance matrices).}
In electrostatics one minimizes $\tfrac12\int_U \epsilon |E|^2\,dV$ subject to boundary potentials on terminal components; the induced terminal charges define the capacitance matrix \cite{jackson_em_1998}.
In magnetostatics one minimizes the magnetic energy subject to prescribed currents/flux constraints; the induced flux linkages define the inductance matrix \cite{jackson_em_1998}.

\subsection{Series and parallel reductions (topological formulation)}
\label{sec:mv-series-parallel}
Series/parallel rules are shadows of two basic Mayer--Vietoris behaviors:
\begin{itemize}
  \item \emph{Direct sums:} when an overlap is homologically trivial in the relevant degree, global state spaces split as direct sums, and quadratic forms add.
  \item \emph{Constraints/kernels:} when gluing imposes matching conditions across an interface, global state spaces are identified with kernels/quotients, and the effective quadratic form is obtained by minimization (Schur complement).
\end{itemize}
Interpreting these induced forms in terminal bases yields the familiar additive/reciprocal series/parallel laws.

\subsection{Circuit reduction laws via Schur complements}
\label{sec:circuit-reductions}
At the lumped level, many textbook reduction laws can be derived from the same variational principle: \emph{add energies of subdevices and eliminate internal variables}.
This is exactly the Schur-complement viewpoint of \S\ref{sec:dtn}.

\begin{proposition}[Series/parallel rules for scalar $R$, $C$, $L$]\label{prop:series-parallel}
Assume ideal, uncoupled elements.
Then the effective parameters satisfy:
\begin{itemize}
  \item \textbf{Resistors.} $R_{\mathrm{series}}=R_1+R_2$ and $G_{\mathrm{parallel}}=G_1+G_2$ where $G_i=R_i^{-1}$.
  \item \textbf{Capacitors.} $C_{\mathrm{parallel}}=C_1+C_2$ and $C_{\mathrm{series}}=(C_1^{-1}+C_2^{-1})^{-1}$.
  \item \textbf{Inductors (no mutual coupling).} $L_{\mathrm{series}}=L_1+L_2$ and $L_{\mathrm{parallel}}=(L_1^{-1}+L_2^{-1})^{-1}$.
\end{itemize}
\end{proposition}

\noindent\emph{Proof sketch.}
For a two-terminal resistor of conductance $G$, the dissipation is a quadratic form in the terminal voltage: $P=\tfrac12 G(V_a-V_b)^2$.
Parallel composition adds dissipation forms directly, giving $G_{\mathrm{parallel}}=G_1+G_2$.
Series composition introduces an internal node potential $V_m$ and one minimizes $P_1(V_a-V_m)+P_2(V_m-V_b)$ over $V_m$; the resulting induced quadratic form has coefficient $G_{\mathrm{series}}=(G_1^{-1}+G_2^{-1})^{-1}$, i.e.\ $R_{\mathrm{series}}=R_1+R_2$.

Capacitors behave identically at the level of energy in voltages: $E=\tfrac12 C(V_a-V_b)^2$, so the same elimination produces $C_{\mathrm{parallel}}=C_1+C_2$ and $C_{\mathrm{series}}=(C_1^{-1}+C_2^{-1})^{-1}$.
Inductors are dual: energy is quadratic in currents, $E=\tfrac12 L I^2$.
Series forces the same current and energies add; parallel splits current and one minimizes $\tfrac12 L_1 I_1^2+\tfrac12 L_2 I_2^2$ subject to $I=I_1+I_2$, yielding $L_{\mathrm{parallel}}=(L_1^{-1}+L_2^{-1})^{-1}$.

\paragraph{Matrix remark.}
For multi-terminal linear passive devices, the same statements hold with scalars replaced by symmetric positive (semi)definite terminal operators and with series/parallel realized as block sums plus Schur complements (the DtN viewpoint).

\subsection{\texorpdfstring{$\Delta$/Y (star--triangle) as Schur complement}{Delta/Y (star--triangle) as Schur complement}}
\label{sec:delta-wye}
Delta--wye equivalence is a canonical example of ``odd gluing'': a $3$-terminal network is simplified by eliminating an internal node.
Historically this is the star--triangle transform \cite{kennelly_1899}; mathematically it is a Schur complement of a discrete DtN/response operator \cite{curtis_ingerman_morrow_1998}.

\paragraph{Operator viewpoint (eliminate the internal node).}
Consider a star (wye) network with terminals $a,b,c$ connected to an internal node $0$ by conductances $g_a,g_b,g_c$.
Writing node potentials as $(V_0,V_a,V_b,V_c)$, the graph Laplacian (nodal admittance matrix) has the block form
\[
L=\begin{pmatrix}
g_a+g_b+g_c & -g_a & -g_b & -g_c\\
-g_a & g_a & 0 & 0\\
-g_b & 0 & g_b & 0\\
-g_c & 0 & 0 & g_c
\end{pmatrix}.
\]
Eliminating $V_0$ (Schur complement) produces an effective $3\times 3$ terminal operator whose off-diagonal entries are $-g_{ab},-g_{bc},-g_{ca}$, i.e.\ an equivalent triangle (delta) network.

\paragraph{Classical conductance formulas.}
Let $g_\Sigma:=g_a+g_b+g_c$.
The equivalent triangle conductances are
\[
g_{ab}=\frac{g_ag_b}{g_\Sigma},\qquad
g_{bc}=\frac{g_bg_c}{g_\Sigma},\qquad
g_{ca}=\frac{g_cg_a}{g_\Sigma}.
\]
(Equivalently, in resistances $R_i=g_i^{-1}$ one obtains $R_{ab}=\frac{R_aR_b+R_bR_c+R_cR_a}{R_c}$ and cyclic permutations.)

\paragraph{Remark (beyond resistors).}
The same Schur-complement algebra applies to any linear passive terminal operator (capacitance matrices, impedance/admittance matrices at fixed frequency), but the resistive case is the cleanest place to see the topology/variational structure explicitly.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  scale=1.05,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  \tikzset{
    tnode/.style={circle,draw,inner sep=2.6pt,line width=0.85pt,fill=white},
    edge/.style={line width=1.1pt},
    elab/.style={fill=white,fill opacity=0.85,text opacity=1,inner sep=1.2pt,rounded corners=1pt}
  }

  % Star (wye) on the left
  \coordinate (O) at (-3.0,0);
  \coordinate (A1) at ($(O)+(90:1.35)$);
  \coordinate (B1) at ($(O)+(-30:1.35)$);
  \coordinate (C1) at ($(O)+(-150:1.35)$);
  \node[tnode] (o) at (O) {$0$};
  \node[tnode] (a1) at (A1) {$a$};
  \node[tnode] (b1) at (B1) {$b$};
  \node[tnode] (c1) at (C1) {$c$};
  \draw[edge] (a1) -- node[midway,elab,left] {$g_a$} (o);
  \draw[edge] (b1) -- node[midway,elab,right] {$g_b$} (o);
  \draw[edge] (c1) -- node[midway,elab,left] {$g_c$} (o);
  \node at (-3.0,2.1) {Y (star)};

  % Arrow: eliminate internal node
  \draw[->,line width=1.0pt]
    (-1.1,0) -- (1.1,0)
    node[midway,above,elab,align=center] {eliminate $0$\\[-2pt]\scriptsize Schur complement};

  % Delta (triangle) on the right
  \coordinate (D) at (3.0,0);
  \coordinate (A2) at ($(D)+(90:1.35)$);
  \coordinate (B2) at ($(D)+(-30:1.35)$);
  \coordinate (C2) at ($(D)+(-150:1.35)$);
  \node[tnode] (a2) at (A2) {$a$};
  \node[tnode] (b2) at (B2) {$b$};
  \node[tnode] (c2) at (C2) {$c$};
  \draw[edge] (a2) -- node[pos=0.55,elab,sloped,above] {$g_{ab}$} (b2);
  \draw[edge] (b2) -- node[midway,elab,below] {$g_{bc}$} (c2);
  \draw[edge] (c2) -- node[pos=0.55,elab,sloped,above] {$g_{ca}$} (a2);
  \node at (3.0,2.1) {$\Delta$ (triangle)};
\end{tikzpicture}
\caption{Delta--wye (star--triangle) equivalence as elimination of an internal node (Schur complement of the response/Laplacian matrix).}
\label{fig:delta-wye}
\end{figure}

\subsection{Optional: a categorical viewpoint}
\label{sec:categorical-measurement}
Many network results become slick once one treats devices not as collections of equations but as maps between boundary data spaces.
In this paper those boundary data spaces are (finite-dimensional) (co)homology groups associated to terminals, and the relevant maps are induced by restriction, duality pairings, and Hodge/constitutive operators.
Gluing corresponds to composition; exactness of Mayer--Vietoris is the mechanism ensuring the composition is consistent.

\section{\texorpdfstring{Alexander duality and degrees of freedom for $C$ and $L$}{Alexander duality and degrees of freedom for C and L}}
\label{sec:alexander-duality}
Alexander duality is the basic bridge between the topology of the conductor region $M\subset S^3$ and the topology of the field domain $U=S^3\setminus\mathrm{int}(M)$.
It is the mechanism by which ``capacitor-type'' disconnectedness of $M$ and ``inductor-type'' linking/loop phenomena in $U$ become the same statement in different degrees.

\paragraph{Proof sketch of \Cref{thm:alexander-rlc}.}
Alexander duality identifies reduced cohomology of the complement with reduced homology of the subset in complementary degree.
Disconnectedness of $M$ is detected by $\widetilde H_0(M)$ and produces degree-$2$ classes in $U$ (capacitive degrees of freedom), while loop classes in $M$ (and tunnels in the complement) produce degree-$1$ classes in $U$ (inductive degrees of freedom).
Once $\epsilon$ and $\mu$ are chosen, these state spaces acquire bilinear forms by integration (Section~\ref{sec:elements-as-pairings}); effective $C/L$ matrices are obtained by DtN reduction (Section~\ref{sec:dtn}).

\subsection{Statement (informal form)}
Let $A\subset S^3$ be a compact, locally contractible subset (in particular, a compact submanifold-with-boundary as in our conductor model).
Then Alexander duality provides natural isomorphisms for reduced (co)homology:
\[
\widetilde H^k(S^3\setminus A;\,R)\;\cong\;\widetilde H_{2-k}(A;\,R),
\]
for all $k$, where $R$ is a coefficient ring/field.
In our setting, $A=M$ (or sometimes a deformation retract of $M$ capturing its topology).

\subsection{Two circuit-relevant consequences}
\paragraph{Capacitors (components of $M$).}
If $M$ has multiple connected components, then $\widetilde H_0(M;R)\neq 0$, hence by Alexander duality $\widetilde H^2(S^3\setminus M;R)\neq 0$.
This is the topological reason that separating $2$-classes in the field domain exist when there are multiple conductors: they support nontrivial electric-flux degrees of freedom (which become a capacitance matrix once $\epsilon$ is chosen).

\paragraph{Inductors (loops/linking in the complement).}
Nontrivial $\widetilde H_1(S^3\setminus M;R)$ (equivalently $H_1(U;R)$) corresponds dually to $\widetilde H^{1}(S^3\setminus M;R)\cong \widetilde H_{1}(M;R)$ in the simplest cases, and more generally is detected by the presence of nontrivial ``tunnels'' in the complement.
This is the topological home for magnetic flux/linking effects that, after choosing $\mu$, become the inductance matrix.

\begin{figure}[H]
\centering
\begin{minipage}{0.48\textwidth}
\centering
\begin{tikzpicture}[
  scale=1.0,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  \tikzset{
    lab/.style={fill=white,fill opacity=0.82,text opacity=1,rounded corners=2pt,inner sep=1.6pt}
  }

  \def\Ru{2.0}
  \def\Rm{0.55}
  \coordinate (Mone) at (-0.9,0.25);
  \coordinate (Mtwo) at (0.9,-0.25);

  % Field domain cross-section
  \shade[ball color=blue!10] (0,0) circle (\Ru);
  \draw[line width=0.95pt] (0,0) circle (\Ru);
  \node[lab,text=blue!60!black] at (0,2.25) {$U$};

  % Two conductor components and terminal subsets
  \shade[ball color=gray!28] (Mone) circle (\Rm);
  \shade[ball color=gray!28] (Mtwo) circle (\Rm);
  \draw[line width=0.95pt] (Mone) circle (\Rm);
  \draw[line width=0.95pt] (Mtwo) circle (\Rm);
  \node at (Mone) {$M_1$};
  \node at (Mtwo) {$M_2$};

  % Terminals (facing each other, for readability)
  \draw[line width=1.5pt,red!80!black]
    (Mone) ++(-35:\Rm) arc (-35:35:\Rm) node[midway,right=2pt,lab,text=red!80!black] {$\Gamma_1$};
  \draw[line width=1.5pt,red!80!black]
    (Mtwo) ++(145:\Rm) arc (145:215:\Rm) node[midway,left=2pt,lab,text=red!80!black] {$\Gamma_2$};

  % Separating surface class (cartoon)
  \path[fill=green!25,opacity=0.35]
    (-0.25,1.60) .. controls (-0.55,0.40) and (0.25,-0.15) .. (0.25,-1.60)
    -- (0.45,-1.60) .. controls (0.75,-0.15) and (-0.05,0.40) .. (-0.45,1.60) -- cycle;
  \draw[line width=1.05pt,green!60!black,dashed]
    (0,1.60) .. controls (-0.20,0.40) and (0.20,-0.15) .. (0,-1.60);
  \node[lab,text=green!60!black,align=center] at (1.05,1.18)
    {$\Sigma$\\[-1pt]\scriptsize $[\Sigma]\in H^2(U)$};

  % A probe path between terminals (relative 1-chain)
  \coordinate (P) at ($(Mone)+(0:\Rm)$);
  \coordinate (Q) at ($(Mtwo)+(180:\Rm)$);
  \draw[line width=1.1pt,->,purple!70!black] (P) to[out=10,in=170]
    node[pos=0.55,lab,text=purple!70!black] {$\gamma$} (Q);
  \fill[purple!70!black] (P) circle (1.15pt);
  \fill[purple!70!black] (Q) circle (1.15pt);
\end{tikzpicture}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\centering
\begin{tikzpicture}[
  scale=1.0,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
  \tikzset{
    lab/.style={fill=white,fill opacity=0.82,text opacity=1,rounded corners=2pt,inner sep=1.6pt}
  }

  % Spanning surface S (drawn as a disk) with boundary loop in M
  \shade[inner color=blue!8,outer color=blue!22,opacity=0.75] (0,0) ellipse (1.6 and 0.85);
  \draw[line width=2.6pt,gray!35] (0,0) ellipse (1.6 and 0.85);
  \draw[line width=1.2pt,gray!75!black] (0,0) ellipse (1.6 and 0.85);
  \node[lab,text=gray!75!black] at (0,1.35) {loop in $M$};
  \node[lab,text=blue!60!black] at (0,-1.35) {surface $S$};

  % Linking cycle gamma (front half solid, back half dashed)
  \draw[line width=1.2pt,red!70!black] (0,1.55) arc (90:-90:0.55 and 1.55);
  \draw[line width=1.2pt,red!70!black,dashed] (0,-1.55) arc (-90:90:0.55 and 1.55);
  \node[lab,text=red!70!black] at (0.78,0.05) {$\gamma$};
  \fill[red!70!black] (0,0) circle (1.6pt);
  \node[lab,text=red!70!black] at (1.0,-0.40) {\scriptsize $\#(S\cap\gamma)\neq 0$};
\end{tikzpicture}
\end{minipage}
\caption{Cartoons of the topological degrees of freedom behind a capacitor (left: a separating surface class in $H^2(U)$ associated to disconnected conductors) and an inductor (right: linking detected by intersection of a cycle with a spanning surface).}
\label{fig:cap-ind-cartoon}
\end{figure}

\subsection{Relative refinements and terminals}
For actual circuit elements we typically distinguish terminal surfaces $\Gamma\subset \partial M$ and impose boundary conditions there.
The correct statements then involve relative (co)homology and Poincar\'e--Lefschetz duality on $U$ (a manifold with boundary), together with excision to pass between $S^3\setminus M$ and $U$.
We will use these relative refinements when deriving capacitance/inductance matrices for multi-terminal devices.

\subsection{Alexander duality as a ``degree-of-freedom counter'' (and when reductions exist)}
\label{sec:alexander-series-parallel}
Alexander duality is most useful here as a clean \emph{counting principle}: it predicts which finite-dimensional topological state spaces \emph{can} exist before any constitutive geometry is chosen.
In particular, it tells you which ``reactive'' degrees of freedom are even available to be turned into $C$ and $L$ matrices by a Hodge/energy construction.

\paragraph{Why this helps with reductions.}
Interconnection operations (series/parallel attachments, identification of terminal components, adding/removing internal nodes) change the topology of the conductor region $M$ in controlled ways, hence change $\widetilde H_0(M)$ and $\widetilde H_1(M)$.
By Alexander duality, this forces corresponding changes in the complement groups (schematically, $H^2(U)$ for capacitive classes and $H^1(U)$ for inductive classes).
Thus topology predicts \emph{which} reductions are meaningful: one can only reduce if the gluing keeps track of the relevant state spaces and their pairings.

For the actual reduction \emph{formulas} (series/parallel and $\Delta$/Y) one needs both Mayer--Vietoris bookkeeping and the variational/DtN mechanism:
see Section~\ref{sec:gluing}, especially \S\ref{sec:mv-series-parallel} and \S\ref{sec:delta-wye}.

\section{Time topology and AC/DC signals}
\label{sec:time-topology}
We model time dependence by taking a product with a $1$-manifold $\mathcal{T}$:
\begin{itemize}
  \item \emph{DC} (transients/step response): $\mathcal{T}=[0,\infty)$ (or $[0,\infty]$ after one-point compactification when using compact-topology arguments).
  \item \emph{AC} (steady periodic forcing): $\mathcal{T}=S^1$ with $\partial \mathcal{T}=\emptyset$.
\end{itemize}
For topological bookkeeping we may replace $[0,\infty)$ by its compactification $[0,\infty]\cong [0,1]$, so the relevant (co)homology of the time factor is elementary:
\[
H_0([0,\infty];R)\cong R,\qquad H_1([0,\infty];R)=0,
\]
and
\[
H_0(S^1;R)\cong R,\qquad H_1(S^1;R)\cong R.
\]
For any space $Y$ used to model spatial/topological degrees of freedom (e.g.\ the circuit complex $X$ or a field-domain pair $(U,\Gamma)$), the product $Y\times \mathcal{T}$ organizes time-dependent degrees of freedom.

\subsection{\texorpdfstring{$Y\times \mathcal{T}$ as a trivial bundle (and what is fixed in the DC case)}{YxTime as a trivial bundle (and what is fixed in the DC case)}}
\label{sec:bundle-viewpoint}
It is useful to regard
\[
\pi\colon Y\times \mathcal{T}\to \mathcal{T}
\]
as the trivial fiber bundle with fiber $Y$.
In this paper we mostly use this bundle in the simplest possible way: a time-dependent field/current/voltage is a section (or family of classes/forms) over $\mathcal{T}$, while the product topology enables K\"unneth and slant-product operations.

\paragraph{DC viewpoint (contractible base).}
When $\mathcal{T}$ is contractible (e.g.\ a time interval for DC/transients), the bundle adds no new topological degree of freedom.
What remains is a family of boundary-value problems on copies of $Y$.
For a connected resistive graph (or an elliptic field problem with suitable boundary conditions), fixing terminal values and a gauge convention determines a unique harmonic/minimizing representative in each fiber.
Equivalently: the terminal data selects a unique state modulo the already-quotiented gauge direction.

\paragraph{Why this matters for future sheaf models.}
This bundle language is a useful staging point for a sheaf-theoretic upgrade: instead of assigning one global linear state space to $Y$, one can assign local solution spaces and gluing conditions over patches (and over time) to accommodate directed/reactive/nonlinear components.
For the present note, however, the product bundle picture is enough to cleanly separate topology of available modes from constitutive dynamics on those modes.

\subsection{\texorpdfstring{Topology-only constraints for DC vs.\ AC (what changes when time is $S^1$)}{Topology-only constraints for DC vs. AC (time circle)}}
\label{sec:dc-vs-ac-topology}
\paragraph{Proof sketch of \Cref{thm:ac-kunneth}.}
Over a field, the K\"unneth theorem gives a direct-sum decomposition
\[
H_k(Y\times \mathcal{T};R)\cong \bigoplus_{i+j=k} H_i(Y;R)\otimes H_j(\mathcal{T};R),
\]
and similarly for cohomology.
Passing from DC ($\mathcal{T}$ contractible) to AC ($\mathcal{T}=S^1$) adds an $H_1(S^1)$ factor coming from the time circle.

Even before choosing a Fourier basis, periodic time enlarges natural state spaces.
For example, if $H^{k}(Y;R)\neq 0$, then K\"unneth produces additional cohomology in degree $k+1$:
\[
H^{k+1}(Y\times S^1;R)\;\supset\; H^{k}(Y;R)\otimes H^{1}(S^1;R).
\]

\paragraph{Engineering intuition: why the $H^1(S^1)$ factor matters.}
In the de Rham picture, a generator of $H^1(S^1;R)$ is represented by the closed $1$-form $dt$ (normalized so $\int_{S^1} dt=1$).
Thus any spatial class $[\alpha]\in H^k(Y;R)$ gives a degree-shifted class
\[
[\alpha\wedge dt]\in H^{k+1}(Y\times S^1;R),
\]
which is precisely the $H^k(Y)\otimes H^1(S^1)$ summand above.
Conceptually, periodic forcing introduces an additional ``phase direction'': spatial degrees of freedom acquire a canonical partner involving the time-circle direction, which is the topological shadow of the in-phase/quadrature splitting that engineers package as complex phasors.
The slant product with $[S^1]$ (Section~\ref{sec:background}) is the corresponding ``integrate over one period'' operation: it extracts time-averaged/period-integral quantities from classes on $Y\times S^1$.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  scale=1.0,
  every node/.style={font=\small},
  >=Stealth,
  line cap=round,
  line join=round
]
		  \tikzset{
		    lab/.style={fill=white,fill opacity=0.82,text opacity=1,rounded corners=2pt,inner sep=1.6pt}
		  }

			  \def\rx{2.4}
			  \def\ry{0.65}
			  \def\h{3.0}
			  \pgfmathsetmacro{\thetaA}{270} % front generator
			  \pgfmathsetmacro{\thetaPatchL}{\thetaA}
			  \pgfmathsetmacro{\thetaPatchR}{\thetaA+22}
			  \pgfmathsetmacro{\thetaPatchM}{(\thetaPatchL+\thetaPatchR)/2}
			  \pgfmathsetmacro{\yPatchB}{0.64*\h}
			  \pgfmathsetmacro{\yPatchT}{0.86*\h}
			  \pgfmathsetmacro{\yPatchM}{(\yPatchB+\yPatchT)/2}

		  % A cylinder cartoon for $Y\times S^1$
		  \shade[left color=gray!10,right color=gray!3] (-\rx,0) rectangle (\rx,\h);
		  \shade[inner color=gray!2,outer color=gray!12] (0,\h) ellipse ({\rx} and {\ry});

	  % Axes/labels
		  \draw[->,line width=0.95pt] (-3.2,0.3) -- (-3.2,2.7) node[midway,left] {$Y$};
		  \node[lab] at (0,-0.95) {$S^1$ (time)};

			  % A small surface patch (cartoon for $\alpha\wedge dt$) on the cylinder wall
			  \path[fill=blue!18,opacity=0.22]
			    ({\rx*cos(\thetaPatchL)},{\yPatchB+\ry*sin(\thetaPatchL)})
			    arc (\thetaPatchL:\thetaPatchR:{\rx} and {\ry})
			    -- ({\rx*cos(\thetaPatchR)},{\yPatchT+\ry*sin(\thetaPatchR)})
			    arc (\thetaPatchR:\thetaPatchL:{\rx} and {\ry})
			    -- cycle;
			  \draw[line width=0.55pt,blue!45!black,opacity=0.75]
			    ({\rx*cos(\thetaPatchL)},{\yPatchB+\ry*sin(\thetaPatchL)})
			    arc (\thetaPatchL:\thetaPatchR:{\rx} and {\ry})
			    -- ({\rx*cos(\thetaPatchR)},{\yPatchT+\ry*sin(\thetaPatchR)})
			    arc (\thetaPatchR:\thetaPatchL:{\rx} and {\ry})
			    -- cycle;
			  \coordinate (Cc) at ({\rx*cos(\thetaPatchM)},{\yPatchM+\ry*sin(\thetaPatchM)});

	  % Cylinder outlines on top (keep the rim crisp over the patch)
	  \draw[line width=0.95pt] (0,\h) ellipse ({\rx} and {\ry});
	  \draw[line width=0.95pt] (-\rx,0) -- (-\rx,\h);
	  \draw[line width=0.95pt] (\rx,0) -- (\rx,\h);
		  \draw[line width=0.95pt,dashed] (\rx,0) arc (0:180:{\rx} and {\ry});
		  \draw[line width=0.95pt] (-\rx,0) arc (180:360:{\rx} and {\ry});

				  % Time-circle direction (dt) on the top rim
				  \draw[->,line width=1.0pt,blue!70!black] ({\rx*cos(300)},{\h+\ry*sin(300)}) arc (300:340:{\rx} and {\ry});
					  \node[lab,text=blue!70!black] at (0,\h+0.80) {$dt$};

					  % A spatial class $\alpha$ (as a vertical generator on the wall)
					  \draw[line width=1.15pt,green!60!black]
					    ({\rx*cos(\thetaA)},{\ry*sin(\thetaA)})
					    --
					    ({\rx*cos(\thetaA)},{\h+\ry*sin(\thetaA)});
					  \node[lab,text=green!60!black,anchor=east,xshift=-3pt]
					    at ({\rx*cos(\thetaA)},{0.45*\h+\ry*sin(\thetaA)}) {$\alpha$};

				  % Patch label (drawn last so it stays readable)
				  \node[lab,text=blue!60!black] at (Cc) {$\alpha\wedge dt$};
				\end{tikzpicture}
\caption{Cartoon of the K\"unneth summand $H^k(Y)\otimes H^1(S^1)$: a spatial class $\alpha$ and the time-circle generator $dt$ combine to produce a degree-shifted class $\alpha\wedge dt$ on $Y\times S^1$.}
\label{fig:yxS1-cartoon}
\end{figure}

\paragraph{Homology of $S^1$ versus Fourier modes.}
Singular homology sees only the constant class and the winding class.
The integer-indexed Fourier modes arise analytically as characters of the circle group (equivalently, one-dimensional irreducible complex representations of $S^1$) or as eigenfunctions of the Laplacian on $S^1$, not as additional homology classes.
The topological role of $S^1$ is to provide a new degree-$1$ time direction that can be tensored with spatial classes; the analytic role is that differentiation on each Fourier mode inserts an $i\omega$ factor.

\paragraph{Representation-theoretic diagonalization (and its limits).}
Time translation makes function spaces on $Y\times S^1$ into representations of the circle group.
If a circuit model is \emph{linear} and \emph{time-translation invariant}, then its response operator commutes with this $S^1$-action, so Fourier/character modes diagonalize the time dependence: each character subspace is invariant, and $\partial_t$ acts there by multiplication by $i\omega$.
This gives a mode-by-mode block decomposition of the response.
However, representation theory does \emph{not} by itself imply the physical system is linear; linearity is a constitutive/interconnection assumption.
What representation theory supplies is the canonical decomposition once linearity and time-translation symmetry are present.

\paragraph{Concrete Fourier/Peter--Weyl form of the decomposition.}
On a Hilbert space such as $L^2(Y\times S^1;\mathbb{C})$, one has the Fourier decomposition
\[
L^2(Y\times S^1;\mathbb{C})
\cong
\widehat{\bigoplus}_{n\in\mathbb{Z}}
L^2(Y;\mathbb{C})\otimes \chi_n,
\qquad
\chi_n(t)=e^{int}.
\]
If $A$ is a linear operator commuting with time translations (for example, a network operator built from time-independent coefficients and polynomial expressions in $\partial_t$), then
\[
A \;\cong\; \widehat{\bigoplus}_{n\in\mathbb{Z}} A_n
\]
with each $A_n$ acting only on the spatial factor.
In particular, if $A=P(\partial_t)$ (or appears in a polynomial matrix pencil in $\partial_t$), then mode-by-mode one substitutes $\partial_t\mapsto in$ (or $i\omega$ after restoring physical units/frequency).

\paragraph{How this fits with K\"unneth.}
K\"unneth decomposes the \emph{topological} state spaces of $Y\times S^1$ into spatial and temporal factors.
Fourier/character analysis decomposes the \emph{analytic} time dependence on the $S^1$ factor.
Together they yield a separation-of-variables picture: topological classes split by degree, and within each such class the linear dynamics splits into independent temporal harmonics.

\paragraph{Harmonic representatives on \texorpdfstring{$Y\times S^1$}{YxS1} (product-metric version).}
Your ``separation of variables for harmonics'' intuition is essentially right, but the precise statement is a tensor-product \emph{direct sum}, not a Cartesian product, and it is most naturally stated for harmonic \emph{forms}.
Assume $Y$ is a compact Riemannian manifold (or a compact manifold-with-boundary with fixed boundary conditions) and equip $Y\times S^1$ with a product metric.
Then the Hodge Laplacian splits on decomposable forms as
\[
\Delta_{Y\times S^1}=\Delta_Y\otimes 1 + 1\otimes \Delta_{S^1},
\]
and the harmonic forms satisfy the degreewise decomposition
\[
\mathcal{H}^k(Y\times S^1)
\cong
\big(\mathcal{H}^k(Y)\otimes \mathcal{H}^0(S^1)\big)
\oplus
\big(\mathcal{H}^{k-1}(Y)\otimes \mathcal{H}^1(S^1)\big),
\]
which is the Hodge-level realization of K\"unneth.
This is the precise version of the heuristic ``harmonics on space--time separate into spatial and temporal pieces.''

\paragraph{Important caution for scalar potentials.}
If one restricts to harmonic \emph{functions} (degree $0$), then on connected compact $S^1$ the only harmonic time factor is the constant mode.
So nonconstant Fourier modes on the circle are \emph{Laplacian eigenfunctions} (characters), not harmonic functions.
In circuit language, one usually solves a spatial harmonic/minimizing problem on each time slice (DC picture) or, after Fourier decomposition, solves a family of spatial boundary-value problems mode-by-mode (AC picture), rather than a single harmonic function on the full product $Y\times S^1$.

\paragraph{Lumped-circuit formulation of the same idea.}
For a linear lumped network on a graph $Y=X$ with terminals fixed, a time-dependent node potential can be expanded as
\[
\phi(t)=\sum_{n\in\mathbb{Z}} \phi_n e^{int}
\]
in the periodic case.
Each coefficient $\phi_n$ lives in the spatial node-potential space (modulo gauge), and the network equations decouple by Fourier mode because the response operator commutes with time translation.
The zero mode is the DC component; the nonzero modes are the AC harmonics.
For instance, an equation of the form
\[
\big(A_0 + A_1\partial_t + A_2\partial_t^2\big)\phi = b
\]
reduces mode-by-mode to
\[
\big(A_0 + in\,A_1 - n^2 A_2\big)\phi_n = b_n,
\]
which is the algebraic origin of impedance/admittance formulas and resonance conditions.
This is the graph-level version of the separation-of-variables picture.

\paragraph{The interval case \texorpdfstring{$[0,\infty)$}{[0,infty)} is different.}
For DC/transients, the relevant symmetry is usually the translation semigroup on $[0,\infty)$ rather than the compact group $S^1$.
There is no discrete character basis analogous to Fourier modes on the circle.
Instead one typically uses Laplace transforms (for causal initial-value problems) or semigroup/spectral methods; depending on boundary conditions one may also use sine/cosine transforms on the half-line.
This is the noncompact/causal analogue of ``diagonalizing time'' and is closely related to the Laplace-transform viewpoint engineers already use.
Topologically this matches the fact that $H_1([0,\infty);R)=0$: there is no new periodic time-cycle class to tensor with spatial classes.

\subsection{\texorpdfstring{Leading/lagging and phase as ``$90^\circ$ rotation'' on the time circle}{Leading/lagging and phase as ``90-degree rotation'' on the time circle}}
\label{sec:leading-lagging}
The most visible dynamical signature of energy storage is phase shift: in sinusoidal steady state, ideal capacitors make current lead voltage, while ideal inductors make current lag voltage.

\paragraph{Bundle viewpoint: time derivatives as vertical derivatives.}
On the product $Y\times S^1$ there are two equally natural projections.
Earlier we emphasized $\pi_{\mathrm{time}}\colon Y\times S^1\to S^1$ (a family of spatial states parametrized by time).
For phase behavior it is often cleaner to use the other projection
\[
\pi_{\mathrm{space}}\colon Y\times S^1\to Y,
\]
viewing the time circle as a (trivial) $S^1$-fiber over each point of $Y$.
Then $\frac{d}{dt}$ is literally differentiation along the $S^1$-fiber direction: it is the infinitesimal generator of the circle action on the time factor.
In a Fourier/character basis, this ``vertical derivative'' becomes multiplication by $i\omega$ (generator of Lie algebra for $\mathfrak{u}(1)$), which is the universal source of quadrature phase shifts.

\paragraph{Time derivatives and $S^1$-modes.}
Working on $\mathcal{T}=S^1$ with coordinate $t$, the operator $\frac{d}{dt}$ acts on Fourier modes $e^{int}$ (exponential map from Lie algebra) by multiplication by $in$.
Thus, mode-by-mode, differentiation corresponds to a $90^\circ$ phase shift.

\paragraph{Ideal capacitor and inductor (mode-by-mode).}
The constitutive laws are
\[
i_C = C\,\frac{dv_C}{dt},
\qquad
v_L = L\,\frac{di_L}{dt}.
\]
On a single Fourier mode of angular frequency $\omega$, these become
\[
i_C = (i\omega C)\,v_C,
\qquad
v_L = (i\omega L)\,i_L,
\]
so $Z_C(\omega)=\frac{1}{i\omega C}$ and $Z_L(\omega)=i\omega L$.

\paragraph{What is (and is not) topological about the $90^\circ$ shift.}
The existence of a capacitive or inductive \emph{state variable} is topological (it is detected by the appropriate (co)homology group being nontrivial).
However, the quadrature phase shift itself does not come from (co)homology alone: it comes from universal interconnection laws introducing a time derivative,
\[
i=\frac{dq}{dt},\qquad v=\frac{d\Phi}{dt},
\]
and hence an $i\omega$ in the frequency domain.
In this sense topology tells us \emph{that} there is a storage coordinate ($q$ or $\Phi$), while the time circle tells us \emph{how} it rotates effort/flow variables by $90^\circ$ on each mode.

\paragraph{Hodge star as a spatial duality (separate from the time rotation).}
The Hodge/constitutive operator $\star_{\mathrm{mat}}$ provides an additional duality on spatial forms and induces positive definite bilinear forms on cohomology by
\[
([\alpha],[\beta])\longmapsto \int_U \alpha\wedge (\star_{\mathrm{mat}}\beta).
\]
This is the mechanism that turns a purely topological degree of freedom into a \emph{numerical} capacitance or inductance, while the $90^\circ$ phase shift comes from the complex structure on $S^1$-modes (the $i\omega$ factor).

\section{Worked examples and optional extensions}
\label{sec:examples}

\subsection{Worked topology examples}
\label{sec:worked-topology-examples}
\subsubsection{Spherical cavity (absolute/relative homology)}
Let $M$ be the spherical shell obtained by removing the unit ball from the ball of radius $2$. Then $\partial M$ has two components, each a $2$-sphere, and the long exact sequence of the pair $(M,\partial M)$ computes the nontrivial groups:
\begin{itemize}
  \item $H_3(M,\partial M;R)\cong R$ generated by the fundamental class $[M]$.
  \item $H_2(\partial M;R)\cong R^2$ generated by the two boundary spheres; likewise $H_0(\partial M;R)\cong R^2$.
  \item $H_2(M;R)\cong R$ represented by a sphere separating the two boundary components.
  \item $H_1(M,\partial M;R)\cong R$ represented by a curve connecting the two boundary components.
\end{itemize}
This is the basic template behind ``capacitor'' behavior: disconnected terminal components and a separating $2$-class paired with a relative $1$-class.

\subsubsection{Linking, Gauss integrals, and mutual inductance (topological viewpoint)}
\label{sec:linking-mutual}
The simplest genuinely $3$-dimensional phenomenon behind inductors is linking.
Topologically, linking is detected by homology/cohomology of the complement and by Alexander duality; analytically, it is computed by the Gauss linking integral.

\paragraph{Linking number.}
Let $\gamma_1,\gamma_2\subset S^3$ be two disjoint, oriented smooth embedded circles.
Their \emph{linking number} $\mathrm{Lk}(\gamma_1,\gamma_2)\in\mathbb{Z}$ can be defined by choosing an oriented spanning surface $S$ with $\partial S=\gamma_1$ and setting
\[
\mathrm{Lk}(\gamma_1,\gamma_2):= \#(S\cap \gamma_2)
\]
as the signed intersection number.

\paragraph{Gauss linking integral.}
If $\gamma_1,\gamma_2$ are parametrized in $\mathbb{R}^3$, the same integer is given by
\[
\mathrm{Lk}(\gamma_1,\gamma_2)
=\frac{1}{4\pi}\int_{\gamma_1}\int_{\gamma_2}
\frac{(x-y)\cdot (dx\times dy)}{\|x-y\|^3}.
\]
This formula is a concrete instance of a cohomological pairing computing an intersection/linking invariant.

\paragraph{Mutual inductance as a geometrically weighted pairing.}
In magnetostatics, the flux through a loop induced by current in another loop can be viewed as a symmetric bilinear form on the relevant ``loop'' state space (typically $H_1(U;R)$ or a relative variant), obtained by composing a topological pairing (linking/intersection, controlled by the complement topology) with a constitutive operator determined by $\mu$ (and the metric).
Topology controls \emph{which} couplings can exist robustly, while geometry/materials determine their numerical values.

\subsection{Hybrid example: a parallel cavity--loop reactive branch}
\label{sec:hybrid-parallel-lc}
This example combines the two basic topological storage mechanisms in one two-terminal device.
Take a parallel connection of
\begin{itemize}
  \item an ``ideal capacitor'' branch modeled topologically by a spherical cavity (inner/outer conductor surfaces, with terminal contacts on the two boundary components), and
  \item an ``ideal inductor'' branch modeled topologically by a loop/link configuration detecting a nontrivial $H_1$ class in the field-domain complement.
\end{itemize}
The right panel of \Cref{fig:parallel-cavity-loop} is intentionally schematic/immersed rather than a faithful embedding in $\mathbb{R}^3$; the point is to display terminal attachments and the relevant topological classes in one picture.

\begin{figure}[H]
\centering
\begin{minipage}{0.48\textwidth}
\centering
\begin{tikzpicture}[
  scale=0.95,
  every node/.style={font=\small},
  line cap=round,
  line join=round,
  >=Stealth
]
  \tikzset{
    tnode/.style={circle,draw,inner sep=2.6pt,line width=0.85pt,fill=white},
    wire/.style={line width=1.1pt},
    lab/.style={fill=white,fill opacity=0.85,text opacity=1,rounded corners=2pt,inner sep=1.4pt}
  }

  \node[tnode] (a) at (-3.2,0) {$a$};
  \node[tnode] (b) at ( 3.2,0) {$b$};

  % rails
  \draw[wire] (a) -- (-2.25,0);
  \draw[wire] (b) -- ( 2.25,0);
  \draw[wire] (-2.25,0) -- (-2.25,1.55);
  \draw[wire] ( 2.25,0) -- ( 2.25,1.55);
  \draw[wire] (-2.25,0) -- (-2.25,-1.55);
  \draw[wire] ( 2.25,0) -- ( 2.25,-1.55);

  % capacitor branch (top)
  \draw[wire] (-2.25,1.55) -- (-0.75,1.55);
  \draw[wire] ( 0.75,1.55) -- ( 2.25,1.55);
  \draw[line width=1.2pt] (-0.28,1.10) -- (-0.28,2.00);
  \draw[line width=1.2pt] ( 0.28,1.10) -- ( 0.28,2.00);
  \node[lab,text=blue!60!black] at (0,2.45) {ideal $C$};
  \node[lab,align=center] at (0,0.95) {\scriptsize spherical cavity\\[-1pt]\scriptsize (topological model)};

  % inductor branch (bottom)
  \draw[wire] (-2.25,-1.55) -- (-1.40,-1.55);
  \draw[wire] ( 1.40,-1.55) -- ( 2.25,-1.55);
  \draw[line width=1.1pt] (-1.40,-1.55) arc (180:0:0.35 and 0.30);
  \draw[line width=1.1pt] (-0.70,-1.55) arc (180:0:0.35 and 0.30);
  \draw[line width=1.1pt] ( 0.00,-1.55) arc (180:0:0.35 and 0.30);
  \draw[line width=1.1pt] ( 0.70,-1.55) arc (180:0:0.35 and 0.30);
  \node[lab,text=green!55!black] at (0,-0.65) {ideal $L$};
  \node[lab,align=center] at (0,-2.35) {\scriptsize circle/link\\[-1pt]\scriptsize (topological model)};

  \node[lab,fill=gray!8] at (0,0) {parallel connection};
\end{tikzpicture}
\end{minipage}
\hfill
\begin{minipage}{0.48\textwidth}
\centering
\begin{tikzpicture}[
  scale=0.95,
  every node/.style={font=\small},
  line cap=round,
  line join=round,
  >=Stealth
]
  \tikzset{
    tnode/.style={circle,draw,inner sep=2.6pt,line width=0.85pt,fill=white},
    wire/.style={line width=1.05pt},
    lab/.style={fill=white,fill opacity=0.85,text opacity=1,rounded corners=2pt,inner sep=1.4pt}
  }

  \node[tnode] (a2) at (-3.2,0) {$a$};
  \node[tnode] (b2) at ( 3.2,0) {$b$};
  \draw[wire] (a2) -- (-2.45,0);
  \draw[wire] (b2) -- ( 2.45,0);
  \draw[wire] (-2.45,0) -- (-2.45,1.55);
  \draw[wire] ( 2.45,0) -- ( 2.45,1.55);
  \draw[wire] (-2.45,0) -- (-2.45,-1.60);
  \draw[wire] ( 2.45,0) -- ( 2.45,-1.60);

  % Top branch: cavity capacitor cartoon (immersed schematic)
  \coordinate (Cc) at (0,1.40);
  \shade[inner color=gray!10,outer color=gray!24,opacity=0.9] (Cc) circle (0.95);
  \shade[inner color=white,outer color=white] (Cc) circle (0.43);
  \draw[line width=0.95pt] (Cc) circle (0.95);
  \draw[line width=0.95pt] (Cc) circle (0.43);
  \draw[wire] (-2.45,1.55) -- (-0.95,1.55) -- ($(Cc)+(-0.95,0)$);
  \draw[wire] ( 2.45,1.55) -- ( 0.25,1.55) -- ($(Cc)+(0.43,0)$);
  \draw[line width=1.5pt,red!75!black] ($(Cc)+(-155:0.95)$) arc (-155:-25:0.95);
  \draw[line width=1.5pt,red!75!black] ($(Cc)+(25:0.43)$) arc (25:155:0.43);
  \draw[line width=0.9pt,green!55!black,dashed] ($(Cc)+(0,0.90)$) -- ($(Cc)+(0,-0.90)$);
  \node[lab,text=red!75!black] at (-1.35,2.45) {$\Gamma_a$};
  \node[lab,text=red!75!black] at (1.55,2.45) {$\Gamma_b$};
  \node[lab,text=green!55!black] at (1.35,1.05) {$\Sigma_C$};

  % Bottom branch: loop/link inductor cartoon
  \coordinate (Lc) at (0,-1.45);
  \draw[wire] (-2.45,-1.60) -- (-1.15,-1.60);
  \draw[wire] ( 2.45,-1.60) -- ( 1.15,-1.60);
  % conductor loop with a small top gap where leads attach
  \draw[line width=1.15pt,gray!70!black] ($(Lc)+(150:0.95)$) arc (150:20:0.95 and 0.58);
  \draw[line width=1.15pt,gray!70!black] ($(Lc)+(-20:0.95)$) arc (-20:-210:0.95 and 0.58);
  \draw[wire] (-1.15,-1.60) -- ($(Lc)+(150:0.95)$);
  \draw[wire] ( 1.15,-1.60) -- ($(Lc)+(30:0.95)$);
  % linked cycle / flux probe
  \draw[line width=1.1pt,blue!65!black] ($(Lc)+(0,1.08)$) arc (90:-90:0.36 and 1.08);
  \draw[line width=1.1pt,blue!65!black,dashed] ($(Lc)+(0,-1.08)$) arc (-90:90:0.36 and 1.08);
  \node[lab,text=gray!70!black] at (0,-0.35) {loop branch};
  \node[lab,text=blue!65!black] at (1.15,-1.45) {$\gamma_L$};

  \node[lab,align=center,fill=gray!8] at (0,0.05)
    {\scriptsize schematic immersed topological realization\\[-1pt]\scriptsize (terminal wires shown explicitly)};
\end{tikzpicture}
\end{minipage}
\caption{A two-terminal parallel LC example combining the two topological storage mechanisms. Left: the circuit interconnection pattern. Right: a schematic/immersed topological realization with a spherical-cavity capacitor branch (top) and a loop/link inductor branch (bottom), both attached to the same terminals.}
\label{fig:parallel-cavity-loop}
\end{figure}

\paragraph{Topology-first analysis of the parallel interconnection.}
Let $U_C$ and $U_L$ denote the field domains for the cavity and loop branches, with terminal subsets glued to the same two terminal components.
Excision removes irrelevant wire geometry away from those terminals, and the gluing occurs only at the terminal interface.
Accordingly, the reactive state space is (schematically) a direct sum of the branch state spaces:
\[
\mathcal{S}_{\mathrm{react}}
\;\cong\;
\Span\{[\Sigma_C]\}\oplus \Span\{[\gamma_L]\},
\]
where $[\Sigma_C]$ denotes the capacitor-type separating class (degree-$2$/relative in the field picture) and $[\gamma_L]$ denotes the inductor-type loop/link class (degree-$1$ in the complement picture).
This is exactly the ``direct sum'' branch of the Mayer--Vietoris dichotomy from \S\ref{sec:mv-series-parallel}; the terminal matching condition is simply that both branches see the same terminal voltage.

\paragraph{Explicit Mayer--Vietoris chase for this parallel LC gluing.}
Let $A$ denote the cavity branch together with a small terminal collar, let $B$ denote the loop branch together with the same collar, and let $I:=A\cap B$ be the overlap (the terminal collar itself).
Topologically, $I$ deformation retracts onto two terminal patches (or, in a lumped caricature, two vertices), so
\[
H^k(I;R)=0\quad (k\ge 1),\qquad H^0(I;R)\cong R^2.
\]
For the reactive degrees (degree $1$ for the inductor-type class and degree $2$ for the capacitor-type class), the Mayer--Vietoris cohomology sequence has segments
\[
H^{m-1}(I;R)\longrightarrow H^m(A\cup B;R)\xrightarrow{j^\ast}
H^m(A;R)\oplus H^m(B;R)\xrightarrow{\rho_A-\rho_B} H^m(I;R),
\]
with $m=1,2$.
For $m=2$, since $H^1(I;R)=H^2(I;R)=0$, exactness gives
\[
H^2(A\cup B;R)\xrightarrow{\;\cong\;} H^2(A;R)\oplus H^2(B;R),
\]
so the capacitor-type degree-$2$ class glues by direct sum.
For $m=1$, one still gets surjectivity
\[
H^1(A\cup B;R)\twoheadrightarrow H^1(A;R)\oplus H^1(B;R),
\]
but the kernel can contain the image of the connecting map $\delta\colon H^0(I;R)\to H^1(A\cup B;R)$; this is exactly the extra ``parallel-loop'' contribution created by gluing along two terminal components.
In the terminal-relative formulation (or after imposing the terminal matching/gauge conditions relevant to the chosen readout), this degree-$0$ contribution is quotiented out, leaving the branch loop class as the intended inductor coordinate.

Concretely, the chase is as follows.
For $m=2$, given branch classes $(\alpha_A,\alpha_B)$, their restrictions to $I$ vanish, so $(\rho_A-\rho_B)(\alpha_A,\alpha_B)=0$ and exactness gives a \emph{unique} global class $\alpha\in H^2(A\cup B;R)$ with $j^\ast(\alpha)=(\alpha_A,\alpha_B)$.
For $m=1$, the same argument gives existence of a global preimage, but uniqueness is only up to the image of $\delta(H^0(I;R))$ unless one passes to the relative/gauge-fixed version.
This is the topological reason the cavity and loop storage coordinates coexist in the parallel device, while also explaining why terminal matching/gauge bookkeeping cannot be ignored in degree $1$.

\paragraph{Where terminal matching appears in the same sequence.}
For degree $0$ (potentials before quotienting by gauge), the same Mayer--Vietoris sequence reads
\[
0\longrightarrow H^0(A\cup B;R)\xrightarrow{j^\ast} H^0(A;R)\oplus H^0(B;R)
\xrightarrow{\rho_A-\rho_B} H^0(I;R)\longrightarrow H^1(A\cup B;R)\longrightarrow \cdots.
\]
Here the map $\rho_A-\rho_B$ is exactly ``difference of terminal restrictions.''
Thus the familiar parallel-circuit condition ``the branches have the same terminal voltage'' is literally the kernel condition in degree $0$.
The degree-$2$ capacitor class glues by direct sum, while the degree-$1$ loop class is controlled by the same sequence together with the connecting map/gauge quotient.

\paragraph{Mode-by-mode AC response (after constitutive data is chosen).}
Once geometry/material data assigns numerical parameters $C$ and $L$, the phasor-domain branch laws at angular frequency $\omega>0$ are
\[
I_C=(i\omega C)V,\qquad I_L=\frac{1}{i\omega L}\,V.
\]
Parallel gluing adds currents, so the effective admittance is
\[
Y_{\mathrm{eff}}(\omega)=\frac{I_C+I_L}{V}
= i\omega C+\frac{1}{i\omega L}.
\]
In particular, the ideal lossless parallel LC branch has an antiresonance at $\omega_0=1/\sqrt{LC}$ where the susceptances cancel.
The topological input is the existence of one capacitor-type and one inductor-type storage coordinate; the constitutive/Hodge input supplies the numerical values $C,L$ and hence the resonance scale.

\paragraph{How this example interfaces with the \texorpdfstring{$Y\times S^1$}{YxS1} picture.}
On $Y\times S^1$, K\"unneth records the extra time-circle factor and Fourier/character decomposition diagonalizes the time dependence.
The same parallel-direct-sum topology persists in each harmonic sector, while $\partial_t$ contributes the mode factor $i\omega$ that distinguishes the capacitor and inductor responses.

\subsection{Worked lumped-circuit examples}
\label{sec:lumped-examples}
These short examples connect the chain/cochain picture to standard computations.

\subsubsection{Single-loop current as a rank-one cycle space}
Let $X$ be any connected graph with one independent loop (first Betti number $1$).
Then $\ker\partial\subset C_1(X;R)$ is one-dimensional, so any KCL-satisfying current is a scalar multiple of a fundamental cycle current.
Constitutive data (resistive/inductive) selects the scalar via an energy/dissipation minimization or via the time-domain ODE for the storage coordinate.

\subsubsection{Two-terminal ``voltage'' as a relative evaluation}
Let $T=\{t_0,t_1\}\subset V$ be two terminals.
Any probe path is a $1$-chain $\gamma$ with $\partial\gamma=t_1-t_0$.
If $v=d\phi$ then $v(\gamma)=\phi(t_1)-\phi(t_0)$, independent of the chosen path.
When $[v]\neq 0$ in $H^1(X;R)$, different probe paths differ by a cycle and the discrepancy is detected by evaluating $[v]$ on that cycle.

\subsubsection{Series/parallel as elimination (Schur complement)}
In a linear resistive network, the power functional is a quadratic form in branch currents/voltages.
Gluing two subnetworks and eliminating internal node potentials is a Schur complement computation on that quadratic form; the resulting induced form on terminal data is the effective resistance operator.
This is the algebraic shadow of the Mayer--Vietoris ``direct sum vs.\ constraint'' dichotomy discussed in \S\ref{sec:mv-series-parallel}.

\subsection{Optional: almost topological degrees of freedom}
\label{sec:almost-topological}
Ideal topological degrees of freedom are defined by exact closedness/exactness and perfect boundary conditions.
Real devices are non-ideal: conductors have finite conductivity, dielectrics leak, and gluing decompositions are only approximately valid.
This optional module records one way to make the ideal picture quantitative.

\paragraph{Functional-analytic setting (minimal).}
In a Maxwell/FEEC-type framework, one works in Sobolev spaces such as $H(\mathrm{curl},U)$ and $H(\mathrm{div},U)$ with boundary conditions encoding terminals \cite{arnold_falk_winther_2010_feec}.
The material inner product $\langle\cdot,\cdot\rangle_{\mathrm{mat}}$ defines a (weighted) Hodge Laplacian $\Delta_{\mathrm{mat}}$ whose kernel corresponds (in ideal lossless settings) to the relevant cohomology classes.

\paragraph{Spectral viewpoint: ``almost'' as a near-kernel.}
Non-idealities perturb exact kernels into small but nonzero eigenvalues.
Thus one may call a mode ``$\varepsilon$-topological'' if its Rayleigh quotient
\[
\lambda(\alpha):=\frac{\langle \Delta_{\mathrm{mat}}\alpha,\alpha\rangle_{\mathrm{mat}}}{\langle \alpha,\alpha\rangle_{\mathrm{mat}}}
\]
is $\le \varepsilon$.
Small eigenvalues correspond to slowly decaying modes (large time constants, high-$Q$ behavior).

\paragraph{Toy example: an ideal capacitor with leakage.}
An ideal capacitor has a conserved charge coordinate $q$ (a ``zero mode'').
Adding a small leakage resistance $R$ in parallel turns conservation into decay:
\[
\frac{dq}{dt} + \frac{1}{RC}\,q = 0.
\]
The eigenvalue $0$ is perturbed to $1/(RC)$, illustrating how an ``exact'' storage mode becomes an ``almost'' mode with a long decay time.

\paragraph{Further directions (brief).}
Possible extensions include sheaf-theoretic models for localized nonlinear components, persistent-homology bookkeeping for switching networks, and more explicit functorial links between Mayer--Vietoris gluing and operator Schur complements.

\bibliographystyle{unsrt}
\bibliography{main}

\end{document}
